---
title: "Practical machine learning"
author: "Ashish Kumar"
date: "August 23, 2014"
output: pdf_document
---


```{r}

# List of packages for session
.packages = c("data.table","caret","ggplot2","knitr","xtable","randomForest","Hmisc","doParallel","foreach")

# Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])

# Load packages into session 
lapply(.packages, require, character.only=TRUE)

# Read train and test data
training_data <- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", na.strings=c("#DIV/0!") )
evaluation_data <- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", na.strings=c("#DIV/0!") )
```


# data cleansing.
```{r}
for(i in c(8:ncol(training_data)-1)) {training_data[,i] = as.numeric(as.character(training_data[,i]))}

for(i in c(8:ncol(evaluation_data)-1)) {evaluation_data[,i] = as.numeric(as.character(evaluation_data[,i]))}

nonmissing_data<- colnames(training_data[colSums(is.na(training_data)) == 0])[-(1:7)]
reduced_data <- training_data[nonmissing_data]
train

idx <- createDataPartition(y=reduced_data$classe, p=0.75, list=FALSE )
train <- reduced_data[idx,]
test <- reduced_data[-idx,]

```{r}
registerDoParallel()
x <- train[-ncol(train)]
y <- train$classe

rf <- foreach(ntree=rep(200, 5), .combine=randomForest::combine, .packages='randomForest') %dopar% {
randomForest(x, y, ntree=ntree) 
}
```
```{r}
predictions1 <- predict(rf, newdata=train)
confusionMatrix(predictions1,train$classe)


predictions2 <- predict(rf, newdata=test)
confusionMatrix(predictions2,test$classe)
```

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}


x <- evaluation_data
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)

answers
setwd("/Users/coursera")
pml_write_files(answers)
```

